---
title: "Tutorial 1 - Introduction to R - Solution"
author: "Jasper Gross"
output:
  html_document:
    df_print: kable
    toc: yes
    toc_depth: '2'
    number_sections: true
    theme: cosmo
---
<style>
body {text-align: justify}
</style>
---

# Preliminaries

This is a short introduction/reminder into the basics of R. We will have a look at the syntax, some basic objects and functions of R. The great advantage of R is that it is free, easy to understand/use and very powerful (on the other hand it is slower than some other software packages). Therefore it is used by many people especially in academia all over the world and its user base is still growing. In fact its huge user base is an great advantage, too. Especially for beginners, since it allows you to write and share programs for specific tasks. It should not surprise you that for many problems you might think of or encounter later on, a solution already exists somewhere (Google is your best friend!). To use these programs of others you will need to install certain so-called packages. For this purpose R offers the function **install.packages("???")** [Note the quotes!]. To be able to use these packages in your session you have to use the command **library(???)**. The installation of a certain package is only needed once, the call to the library every time you restart R! 

To make things work you have to first install the four packages and activate them. To do so press the small green run button in the grey box below:

```{r eval=FALSE}
install.packages("rmarkdown")
install.packages("formatR")
install.packages("caTools")
install.packages("knitr")
library(rmarkdown)
library(formatR)
library(caTools)
library(knitr)
```

Especially in the beginning a very useful function is the help. You can open it in R by typing **help(???)**.

```{r warning=FALSE, error=TRUE}
# Call the help page of the command help
help(help)
```

Note that the hash tag (\#) can be used in R scripts to make comments. Comments are just for the reader of the code and will not be executed by R. A good program should include comments where the code is explained. This is not only important for other users of your code but also for yourself because projects might become fairly large or you might work a long time on a given project.


# Numbers, vectors and basic functions
Of course R could be used as an (advanced) pocket calculator by using the operator **"+"**, **"-"**, **"$\mathbf{\cdot}$"** or **"/"**. You can store the results in variables. The name of the variable is always on the left, the statement you would like to save on the right. We use **"="** or **"<-"** to assign something to a variable name. Note that there are certain rules for variable names: It has to begin with a letter, special characters are (mostly) not allowed and certain names as *TRUE* or *FALSE* are reserved for R commands or functions. If you use the same name for a variable twice you overwrite the content of the former.  

Calculate the square of 5 and save the result in the variable x. What is x? Now assign the value 7 to x. What is x now?

```{r   warning=FALSE, error=TRUE}
x = 5^2
x
x = 7
x
```

You can have a look at the content of any variable by calling the name - as we did for x. Very often you will deal with (large) data sets. The most basic R object to save data - which is called data structure - is a vector. A vector is easy to construct via the function **c(???)**. 

```{r   warning=FALSE, error=TRUE}
vec = c(3, 5, 7, 9, 11, 13, 15)
vec
```

Sometimes this takes time to type out therefore another useful function to create a vector based on a certain pattern is **seq(???)**. Any function in R has a name - e.g. **"seq"** - and round brackets **"()"**, pretty much like in mathematics. The *???* are called arguments. If you have a look at the help page of the function **seq()**

```{r   warning=FALSE, error=TRUE}
help(seq)
```

you will notice that this is structured in different parts which are very typical for (most) help pages of functions. First a description what the function is intended to do, the explanation of the syntax (Usage) followed by a list of potential arguments and at the very end you find very often some examples to illustrate the usage. 

Try to create the same vector as in the previous part using the function **seq()**.

```{r   warning=FALSE, error=TRUE}
vec2 = seq(3, 15, 2)
vec2
```

Most arguments of any function have a name. It is good practice to specify the name of the argument you would like to use. The order of the arguments matter, e.g. for the function **seq()** the first argument *from* is the starting point, the second one *to* the end point of the sequence. Arguments are always separated by a comma. You will also notice that the argument *from* has a so called default value (which is 1 in this case). This implies if you do not specify this argument it will take the value 1 (this is sometimes very useful, especially if you write your own functions). 

Create a sequence of integers starting at 1 ending at 10 by specifying all arguments (with names). 

```{r   warning=FALSE, error=TRUE}
s = seq(from = 1, to = 10, length.out = 10)
s
```


A very important difference to some other programming languages is that R will manipulate (numeric) vectors element wise. For example the scalar product of two vectors a and b of the same length *<a,b>* is typically *ab* but not in R!

Construct two numeric vectors a and b with the same number of elements (n = 4) and multiply them

```{r   warning=FALSE, error=TRUE}
a = c(1,2,3,4)
b = c(5,6,7,8)
a*b           # elementwise
a%*%b         # scalar product
```

In some applications this is very useful in other applications you really have to think hard how to translate element wise manipulations of a vector to your desired outcome. If you multiply (divide, ...) vectors of different length the shorter vector will be recycled automatically. Note that if the length of the shorter is an exact multiple of the larger vector you will not even get an error message, otherwise only a warning.  

Sometimes you may be interested in certain subsets of a given vector. In R this is possible by specifying the name and the index number of the elements you are interested in, e.g. *vec[1:3]* which extracts the first three elements of the vector vec. 

How can you obtain the elements 2 to 5 and 7 of the vector vec?

```{r   warning=FALSE, error=TRUE}
vec[1:3]
vec[c(2:5,7)]
```

Notice that if you would like to obtain elements of a vector (or later any data structure) you have to use rectangular brackets **"[]"**, if you would like to call a function always round brackets **"()"**. You can exclude certain elements of a vector with a minus, e.g. *vec[-(2:5)]*.


# Loops
Whenever something has to be repetitively executed the most basic procedure to solve such a task is a so called *loop*. Loops are just designed to repeatedly execute (similar) expressions. In R two basic types of loops exist: 1. the so called *"for"* loop and 2. the *"while"* loop (to be more precise there exist a third option: repeat loop, but this behaves pretty much like the while loop). The for loop is easier to handle, therefore we begin with that. You might have a look at the help page. The syntax for a for loop is quiet similar to that for 'if' statements. 

Suppose that we want to calculate the sum of the variable s from earlier without using the *sum()* function. 

```{r   warning=FALSE, error=TRUE}
sum = 0
for(i in 1:length(s)){
  sum=sum+s[i]
}
sum
```
  
If you have a look at the list of variables, the variable sum has the value 42 which is indeed the sum of all elements of x (you may use *sum()* to verify this result). We need at first some condition based on an index (in our example called i) how often (and for which 'values') something has to be calculated. Since we want to calculate the sum of x we start with the first element x[1] and end with the last one, therefore we used the *length()* function. After each step the index is increased by 1. The calculation looks a little bit strange from a mathematical perspective because the equality *sum = sum +x[i]* almost never holds. But remember that R calculates the right hand side (in the first step 0 + 1) and save the result in the already existing variable sum. This procedure is very often useful. Start with an variable set to 0 and calculate something based on this variable. Save the result in the variable again.  

The second type of loops are the so called while loops. The syntax is similar to the for loop with a main difference. The statement within the **{}**-brackets is executed as long as the condition in the round brackets **()** is true! This yields a major problem because the condition may be always true, therefore the while loop will never stop (infinite loop)!!! 

Again calculate the sum of s:

```{r   warning=FALSE, error=TRUE}
i = 1
sum = 0
while(i<=length(s)){
  sum = sum + s[i]
  i = i+1
}
sum
```

To ensure that the while loop will stop we have to increase the index i within the loop.  

Loops are relatively easy to use, unfortunately they are relatively slow. If you are running for example larger simulations it might pay off to use vectorized commands or functions instead of loops to save a lot of computation time.


# Writing functions

R is a functional language which means almost everything is a function (besides data of course). It is highly recommended to use this mind-set, too. Thus if you have to implement a new procedure for a given task, analyzing data or running a simulation you might write your own function who does exactly this. Typically this is not one big function but a function that makes use of other functions (which you also have written for this purpose). The goal for almost any larger project is to disaggregate this project into smaller parts that can be handled separately (maybe with some dependency between those) and write a function for each sub-part.  

Now let's write our first function. You need several things to define a function. A name, some arguments or variables that the function may use and of course a procedure. The syntax is as follows:

```{r   warning=FALSE, error=TRUE}
name = function(argument1, argument2, argument3 = a,...){
  procedure
  return(result)
}
```

Note that the third argument has a default value, i.e. a value which is used whenever nothing is specified for this argument by the user. 

Try writing a function that calculates the mean of any numerical vector. Use the functions **sum()** and **length()** for this purpose. Test this function with the vector a you defined earlier.

```{r   warning=FALSE, error=TRUE}
our_mean = function(x){
  result = sum(x)/length(x)
  return(result)
}

our_mean(a)
identical(our_mean(a), mean(a))
```

We calculated the mean by summing up all elements of the variable a and divide the sum by the number of elements. The word function is the signal word for R that a function is defined. Note that we used only one argument x. The variable x is called a local variable because it exists only in our function but not in any other environment (have a look at the upper right)! Local variables are "forgotten" whenever the procedure was executed. Sometimes one uses the same names for local and global variables or arguments. You need to be able to distinguish between them. In fact whenever you write a program you will not use the console. Instead use scripts! Scripts are text files which will be executed by pressing the run button (you have to save them first). Alternatively you may use such a notebook file which combines to some extend text, R commands and results/output.  


# Working with a data set

Task: Load the package **'MASS'** and the dataset named **'Boston'**. Save this set in the variable 'Bost'.

```{r   warning=FALSE, error=TRUE}
library(MASS)
Bost = Boston
```

Hint: In the case you want to work with the dataset you already had in your own PC, look at the 'Environment' tab from the top-right window, then click to 'Import Dataset' and follow the instructions.

Task: Have a look at the Boston dataset

```{r  eval=FALSE, warning=FALSE, error=TRUE}
View(Bost)
```

Task: Get the variable names of all variables in the Boston dataset

```{r   warning=FALSE, error=TRUE}
names(Bost)
```

Task: Compute descriptive statistics and frequencies of the Boston dataset (meaning compute the minimum and maximum values, mode, median, etc of each variable in the dataset)

```{r warning=FALSE, error=TRUE}
summary(Bost)
```

Task: Extract the third variable in the Boston dataset.

```{r warning=FALSE, error=TRUE}
Bost[,3]
```

Task: Take all values from the first to the 90th observations of the 7th variable in the dataset

```{r warning=FALSE, error=TRUE}
Bost[1:90,7]
```

Task: Create a new dataset named 'sub_Bost_1' including variables 1 to 4 and the last three variables

```{r warning=FALSE, error=TRUE}
sub_Bost_1 = Bost[,c(1:4,12:14)]
```

Hint: The task asks us to extract the dataset into the subdataset including variables number 1, 2, 3, 4, 12, 13 and 14. So, we have to indicate these indices to R. 

Task: Create a new dataset name 'sub_Bost_2' including all variables of the dataset Boston except the third variable

```{r warning=FALSE, error=TRUE}
sub_Bost_2 = Bost[,-3]
```

Task: Find the mean, variance, range, 25% and 75% quantile and draw the box plot of the first variable in the Boston dataset.

```{r warning=FALSE, error=TRUE}
mean(Bost[,1])
var(Bost[,1])
range(Bost[,1])
quantile(Bost[,1])
boxplot(Bost[,1])
```

Task: Repeat the procedure by using the name of the variable 'crim' instead.

```{r   warning=FALSE, error=TRUE}
mean(Bost$crim)
var(Bost$crim)
range(Bost$crim)
quantile(Bost$crim)
boxplot(Bost$crim)
```

Task: You suspect that the variables 'tax' and 'indus' are correlated. Display the correlation coefficient matrix to find out whether this may be the case.

```{r warning=FALSE, error=TRUE}
cor(Bost$tax, Bost$indus)
```

Task: Draw a scatter plot to show graphically this relationship, label the x-axis by 'industry', the y-axis by 'tax'.

```{r warning=FALSE, error=TRUE}
plot(Bost$indus, Bost$tax, xlab = 'industry', ylab = 'tax')
```

Task: Regress 'tax' on 'indus' by OLS with **lm(???)**. Assign the result to a new variable named 'result'.

```{r warning=FALSE, error=TRUE}
result = lm(tax ~ indus, data = Bost)
```

Task: Summarize the result of the above regression

```{r warning=FALSE, error=TRUE}
summary(result)
```

Task: Plot the estimated line and the scatter plot in the same figure

Instruction:
1. Create x and y values of the regression curve
2. Draw the scatter plot between 'indus' and 'tax'
3. Draw a line plot between x and y in the same figure, use command 'lines'

```{r warning=FALSE, error=TRUE}
x = seq(min(Bost[,"indus"]), max(Bost[,"indus"]), by = 0.05)
y = result$coefficients[1] + result$coefficients[2]*x
plot(Bost[,'indus'], Bost[,'tax'], xlim = range(Bost[,'indus']), ylim = range(Bost[,'tax']), xlab = 'industry', ylab = 'tax')
lines(x, y, lwd = 2, col = "blue")
```


# Simulations
First, we have to draw a sample of some random numbers. We will start with 10 values of the standard normal distribution.
(By setting a **seed** you'll get always the same random numbers. This is useful for comparing results.)

```{r warning=FALSE, error=TRUE}
set.seed(666)                   
x = rnorm(10, mean = 0, sd = 1)
x
```

Assume we are interested in the distribution of the t-test statistic. Set up a model - with 100 observations - which we can estimate via OLS. Define x as a uniform and u as a normal distributed random variable.

```{r warning=FALSE, error=TRUE}
x = runif(100)
u = rnorm(100, sd = 2)
y = 1 + 2*x + u
```

Now estimate this model via OLS:

```{r warning=FALSE, error=TRUE}
estim = lm(y~x)
coeff = estim$coefficients
covariance = vcov(estim)
```

Construct the t-test statistic for testing the second parameter in our model whether it is 2: 

```{r warning=FALSE, error=TRUE}
t = (coeff[2]-2)/sqrt(covariance[2,2])
```

Unfortunately this is only one value for the test statistic. If we would like to obtain the empirical distribution, we have to repeat this procedure a couple of times. To do this write a function that simulates and estimates a model and returns the test statistic (for the second parameter):

```{r warning=FALSE, error=TRUE}
statistic = function(){
  x = runif(100)
  u = rnorm(100, sd = 2)
  y = 1 + 2*x + u
  estim = lm(y~x)
  coeff = estim$coefficients
  covariance = vcov(estim)
  t = (coeff[2]-2)/sqrt(covariance[2,2])
  return(t)
}
```

Now we can use this function for the construction of a sequence of test statistics

```{r warning=FALSE, error=TRUE}
set.seed(666)
N_MC = 1000 # number of replications in the simulation
simulation = replicate(N_MC, statistic())
```

We can use this sequence to plot the empirical distribution and compare this against the theoretical distribution:

```{r warning=FALSE, error=TRUE}
# empirical:
hist(simulation, breaks = 40, probability = T, xlab = "t statistic")

# theoretical:
x = seq(-5, 5, length.out = 10000)
y = dnorm(x, sd = 1)
lines(x, y, col = "green", lwd = 2)
box()
```

