---
title: "Inferential Statistics Home Assignment 2024"
author: "Jasper Bär & Lukas Baumann"
output: pdf_document
---

# Procedure

-   Try to solve this home assignment and answer all questions in this file by filling the boxes below each exercise.
-   You can and earn up to 6 bonus points for the final exam.
-   You can work in groups of up to 5 students.
-   Hand in your solution (as a Rmd file!) via [e-mail] ([l.baumann\@stat-econ.uni-kiel.de](mailto:l.baumann@stat-econ.uni-kiel.de){.email}) until Friday, 12.07.2024, 23:59:59.
-   Make sure to include all names and stu-numbers in your mail and in the header of this file!
-   We won't evaluate your assignment if your code is not running!
-   Add some comments to your code lines so people can understand your code more easily.
-   We might ask you to explain your code if we suspect that you did not write it on your own.

!!! Enter your group name and stu-numbers here !!!

1.  First name last name, stuXXXXX
2.  First name last name, stuXXXXX
3.  First name last name, stuXXXXX
4.  First name last name, stuXXXXX
5.  Robert Hennings, stu236320

!!! Enter your group name and stu-numbers here !!!

### Task 1 \### (3 points)

Consider the sequence of functions $f_n(x) = \frac{n x}{1 + n^2 x^2}$ for $n = 1, 2, 3, \ldots$ and $x \in [0,1]$. Your task is to explore whether this sequence converges to the function $f(x) = 0$ as n increases.

Hint: You might find sapply() helpful for this task. However, there are different ways to find a solution to this problem.

# a)

Write a function in R that computes $f_n(x)$ for given $n$ and $x$. Furthermore, write the target function $f(x)=0$.

```{r}
# set up the comparing function
f_n_x <-  function(x, n){
  # check for function types and number ranges
  if((x>=0 & x<=1)==F){
    print("x has to be in the range [0, 1]")
  }
  # set up helper function to check for a whole number as described in docs
  is.wholenumber <-
    function(x, tol = .Machine$double.eps^0.5)  abs(x - round(x)) < tol
  if (is.wholenumber(n)==F){
    return("n must be a positive integer!")
  }
  result <- (n*x) / (1 + n^2 * x^2)
  return(result)
}
# Test the first function
result <- f_n_x(x=0.9, n=2)
print(result)

# set up the target function
f_x <- function(x){
  # simply returns 0
  len <- length(x)
  return(rep(0, len))
}
# Test the second function
result <- f_x(x=0.9)
print(result)
```

Choose a few fixed points $x = 0.1, 0.3, 0.5, 0.7, 0.9, 1.0$ and values for $n = c(1:100)$ create a single plot showing the values of $f_n(x)$ for the different fixed points as $n$ increases. Add a meaningful title, labels for the y- and x-axis as well as a legend to the plot.

For which type of convergence are you testing with this task? Does convergence for this concept hold?

```{r}
x <- c(0.1, 0.3, 0.5, 0.7, 0.9, 1.0)
n <- floor(seq(from=1, to=100, length.out=length(x))) # length can be set arbitrarily

get_result_df <- function(x, n){
  # create an empty matrix, convert to data.frame for data storage
  result_df <- data.frame(data=matrix(data=0, nrow=length(n), ncol=length(x)),
                          row.names = as.character(n))
  colnames(result_df) <- as.character(x)
  # iterate over the test size n, holding it fixed
  # and then loop through all the given x-values 
  for(n_ in n){
    # iterate over the values with fixed n
    for(x_ in x){
      result_f_n_x <- f_n_x(x=x_, n=n_)
      # save the result at the correct spot in the matrix
      result_df[as.character(n_), as.character(x_)] <- result_f_n_x
    }
  }
  return(result_df)
}

# create the single plot
# plot settings
result_df <- get_result_df(x=x, n=n)
col_vals <- c("blue", "green", "black", "red", "yellow", "orange") # must equal the length of n
main <- "Test for convergence for different values x in [0, 1] and sizes n in 1:100"

get_comparison_plot <- function(result_df, x, n, col_vals, main){
  for(ind in 1:dim(result_df)[1]){
    row <- rownames(result_df)[ind]
    col <- col_vals[ind]
    if(row==rownames(result_df)[1]){
      # plot the initial figure to which lines will be added afterwards
      plot(x=x,
           y=result_df[as.character(row), ],
           xlim=c(min(x), max(x)),
           ylim=c(min(result_df), max(result_df)),
           col=col,
           xlab="X-values",
           ylab="Function-values: f_x_n",
           main=main,
           cex.main=0.8,
           type="l"
      )  
    }else{
      lines(x=x,
            y=result_df[as.character(row), ],
            col=col)
    }
  }
  legend("topright",
         title = "Size",
         legend = n,
         col = col_vals,
         lty = c(1, 1, 1),
         lwd=c(1, 1, 1)
  )
}

get_comparison_plot(result_df=result_df, x=x, n=n, col_vals=col_vals, main=main)

# Answer here
# convergence in distribution, in probability
```

# b)

Now you are supposed to create a sequence of plots showing $f_n(x)$ for $n = 10, 20, 50, 100, 200$ over the interval $x \in [0, 1]$ in one single figure. Add the target function $f(x) = 0$ to each plot. Add a meaningful title, labels for the y- and x-axis as well as a legend to the plots.

For which type of convergence are you testing with this task? Does convergence for this concept hold?

Hint: You may use par(mfrow=c(3,2)) to create the figure.

```{r}
# plot settings
n <- c(10, 20, 50, 100, 200)
x <- seq(from=0, to=1, length.out=5)
result_df <- get_result_df(x=x, n=n)
par(mfrow=c(3,2))
for(n_ in n){
  # only provide partial result_df to get one plot at a time for one n
  # so only provide the respective row of the result_df
  result_df_n_ <- result_df[as.character(n_), ]
  main <- sprintf("Convergence for size n: %s", n_)
  get_comparison_plot(result_df=result_df_n_, x=x, n=n_, col_vals=col_vals, main=main)
  lines(x=x, y=f_x(x), col="red")
  legend("topright",
         title = "Size",
         legend = c(n_, "f(x)"),
         col = c(col_vals[1], "red"),
         lty = c(1, 1),
         lwd=c(1, 1)
  )
}
par(mfrow=c(1,1))

# Answer here:
# 
```

Choose an $\epsilon > 0$, for instance, $\epsilon = 0.1$. Plot $|f_n(x) - f(x)|$ for $n = 1, 2, 5, 10, 20$ and show that there are always some values of $x$ in $[0, 1]$ such that $|f_n(x) - f(x)| > \epsilon$ to support your argument in (b).

What do the plots show and what can we conclude from this observation?

```{r}
# plot settings
epsilon = 0.1
n <- c(1, 2, 5, 10, 20)
result_df <- get_result_df(x=x, n=n)

par(mfrow=c(3,2))
for(n_ in n){
  result_df_n_ <- result_df[as.character(n_), ]
  # Take the absolute difference between function and target function
  abs_diff <- abs(result_df_n_ - f_x(x))
  
  main <- sprintf("Convergence for size n: %s and epsilon: %s", n_, epsilon)
  plot(x=x, y=abs_diff, col="blue", main=main,
       type="l",
       xlim=c(min(x), max(x)),
       ylim=c(min(abs_diff), max(abs_diff)))
  lines(x=x, y=f_x(x), col="red")
  abline(a=epsilon, b=0, col="orange")
  legend("topright",
         title = "",
         legend = c("Abs. Diff.", "f(x)", "e"),
         col = c(col_vals[1], "red", "orange"),
         lty = c(1, 1, 1),
         lwd=c(1, 1, 1)
  )
}
par(mfrow=c(1,1))
#Answer here:
#
```

### Task 2 \### (3 Points)

Calculate Power for a permutation test

For this problem, the following simple regression model is given $y_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \varepsilon_i$, where $\quad i = 1, 2, \ldots, n$, $X_1 \sim \mathcal{U}(1,10)$, $X_2 \overset{\text{iid}}{\sim} \text{Bernoulli}(p)$, and $\varepsilon \sim \mathcal{N}(0,1)$. Further, you can assume that $y_i = 1 + 0.2 x_{1i} + 0.1 x_{2i} + \varepsilon_i$.

Whenever using random numbers please use seed = 100.

# a) Create Data

Write two functions. The first should set up the model. The second should estimates the coefficient $\beta_1$ and its t-value. You may use the $lm()$ function for this step.

```{r}
# settings
set_seed <- T # set seed or not
seed <- 100 # if seed set 100 acc. to task
n <- 100 # sample size
p <- 0.5 # Binom dist
beta_0 <- 1
beta_1 <- 0.2
beta_2 <- 0.1

# ###################################function 1) model setup and data generation
get_simulated_data <- function(set_seed, seed, n, p, beta_0, beta_1, beta_2){
  if(set_seed==T){
    set.seed(seed)  
  }
  x1 <- runif(n=n, min=1, max=10) # specify the data as given in the task
  x2 <- rbinom(n=n, size=1, prob=p) # specify the data as given in the task
  epsilon <- rnorm(n=n, mean=0, sd=1) # standard normal distributed error terms
  # estimate the data with the given parameter values for the coefficients
  y <- beta_0 + beta_1 * x1 + beta_2 * x2 + epsilon
  model_data <- data.frame(X1=x1, X2=x2, Y=y)
  return(model_data)
}
# Test the first function
sim_data <- get_simulated_data(set_seed, seed, n, p, beta_0, beta_1, beta_2)
print(sim_data)
# ##################################################function 2) model estimation
model_data <- sim_data
estimate_coeff <- "X1"

get_estimates <- function(model_data, estimate_coeff){
  mod <- lm(model_data$Y ~ model_data$X1 + model_data$X2) # estimate via lm()
  # extract the wanted coefficients and statistics
  estimate <- summary.lm(mod)$coefficients[sprintf("model_data$%s", estimate_coeff), c("Estimate", "t value")]
  return(estimate)
}
# Test the second function
estimates <- get_estimates(model_data=model_data, estimate_coeff=estimate_coeff)
print(estimates)
#  Estimate   t value 
# 0.1788579 5.2328356 

```

# b) Perform a permutation test

Perform a permutation test for the effect of $X_1$ on $Y$. First, generate data with the function you just created. Then, randomly sample $Y$ 30 times and fit the model for each sampled dataset. For each sampled dataset, calculate the coefficient for $\beta_1$ and its t-value. Calculate the p-value as the proportion of times the absolute t-values from the sampled data are greater than or equal to the absolute t-value from the original data. Repeat these steps 100 times and calculate the share of rejections at the 5% level. Perform this process for different sample sizes, starting at 50 and increasing by 50 until reaching 500.

```{r}
# task settings
sim_size <- 100
permutation_size <- 30
rejection_level <- 0.05
n <- seq(from=50, to=500, by=50)
# for later saving the results
results_df <- data.frame(SizeSim = n, matrix(data=NA, nrow=length(n), ncol=sim_size))
colnames(results_df) <- c("SizeSim", 1:sim_size)
# save the single p-vals
p_val_df <- data.frame(SizeSim = n, matrix(data=NA, nrow=length(n), ncol=sim_size))
colnames(p_val_df) <- c("SizeSim", 1:sim_size)
# perform the permutation test
for(n_ in n){
  # we hold the sample size n constant and iterate over the sim_size of 100
  # where we randomly sample Y 30 times in each of the 100 runs
  for(sim in 1:sim_size){
    # loop through different sample sizes contained in the vector n
    # and generate data
    # set the seed to FALSE
    set_seed <- F
    seed <- 100
    p <- 0.8
    beta_0 <- 1
    beta_1 <- 0.2
    beta_2 <- 0.1
    sim_data <- get_simulated_data(set_seed=set_seed, seed=seed, n=n_, p=p,
                                   beta_0=beta_0, beta_1=beta_1, beta_2=beta_2)
    # First estimate the original model
    model_data <- sim_data
    estimate_coeff <- "X1"
    estimates_og <- get_estimates(model_data=model_data, estimate_coeff=estimate_coeff)
    # then conduct the permutation/sampling of Y mixing up its theoretical correct order
    # with which we estimated the original model above, we permute 30 times
    # resulting in 30 estimated values we store accordingly in the following data.frame
    estimates_df_n <- data.frame(Num_Permutation=1:permutation_size, Estimate=NA, t_val=NA)
    for(i in 1:permutation_size){ # 30 times shuffling
      sim_data$Y <- sample(sim_data$Y) # shuffle the Y values, leave Xs as is
      model_data <- sim_data
      estimates_n <- get_estimates(model_data=model_data, estimate_coeff=estimate_coeff)
      # save the estimates
      estimates_df_n[i, 2:dim(estimates_df_n)[2]] <- estimates_n
    }
    # Calculate the p value for the whole n_, from the 30 values we receive 2
    num_true <- sum(abs(estimates_df_n$t_val) >= abs(estimates_og["t value"]))
    # calculate the share out of the 30 values
    p_val <- (num_true / permutation_size)
    p_val_df[p_val_df$SizeSim == n_, sim+1] <- p_val
    # Save values in the estimates df
    # estimates_df[estimates_df$SizeSim == n_, 2:dim(estimates_df)[2]] <- c(num_true, p_val)
    # compute whether we have to reject based on the p-value resulting
    # out of the 30 permutations
    # compare the p-value with the given rejection level
    decision <- ifelse(p_val<rejection_level, "Reject", "NotReject")
    # save the result in the master df
    print(sprintf("Sample size: %s and simulation run: %s and decision: %s", n_, sim, decision))
    results_df[results_df$SizeSim == n_, sim+1] <- decision
  }
}
# Calculate the share of rejections
num_reject <- length(which(results_df=="Reject"))
tota_sims <- length(n) * sim_size
sprintf("The H0 hypothesis is rejected: %s times out of a total of: %s simulations what equals a relative share of: %s percent", num_reject, tota_sims, round(num_reject/tota_sims)*100)

```

# c) Compare test results

Repeat the steps from task b), but set $beta_1 = 0$ for the data generation. Plot the results for the different samples sizes in the same plot as the results for taks b) and compare. Interpret the plot.

```{r}
# iterate over the sample sizes and plot the p value results of simulations from
# a) and b)
for(i in 1:length(n)){
  n_ <- n[i]
  col_ <- col_vals[i]
  if(n_ == n[1]){
    plot(x=colnames(p_val_df),
         y=p_val_df[p_val_df$SizeSim==n_, ],
         type="l",
         ylim=c(0, max(p_val_df[, 2:dim(p_val_df)[2]])),
         col=col_,
         xlab=sprintf("Number of permutations that were carried out %s times each run", sim_size),
         ylab="Estimated P-vals",
         main=sprintf("Permutation Test for sim_size: %s, permutation_size: %s, rejection_level: %s, betas: %s",
                      sim_size, permutation_size, rejection_level, beta_0),
         cex.main=0.8
           )
  }else{
    lines(x=colnames(p_val_df), y=p_val_df[p_val_df$SizeSim==n_, ],
          col=col_)
  }
}
legend("topright",
       title = "Sim Size",
       legend = n,
       col = col_vals,
       lty = rep(1, length(col_vals)),
       lwd=rep(1, length(col_vals))
)

```
