---
title: "Inferential Statistics Home Assignment 2024"
author: "Jasper Bär & Lukas Baumann"
output: pdf_document
---

# Procedure

-   Try to solve this home assignment and answer all questions in this file by filling the boxes below each exercise.
-   You can and earn up to 6 bonus points for the final exam.
-   You can work in groups of up to 5 students.
-   Hand in your solution (as a Rmd file!) via [e-mail] ([l.baumann\@stat-econ.uni-kiel.de](mailto:l.baumann@stat-econ.uni-kiel.de){.email}) until Friday, 12.07.2024, 23:59:59.
-   Make sure to include all names and stu-numbers in your mail and in the header of this file!
-   We won't evaluate your assignment if your code is not running!
-   Add some comments to your code lines so people can understand your code more easily.
-   We might ask you to explain your code if we suspect that you did not write it on your own.

!!! Enter your group name and stu-numbers here !!!

1.  First name last name, stu245231
2.  First name last name, stuXXXXX
3.  First name last name, stuXXXXX
4.  First name last name, stuXXXXX
5.  First name last name, stuXXXXX

!!! Enter your group name and stu-numbers here !!!

### Task 1 \### (3 points)

Consider the sequence of functions $f_n(x) = \frac{n x}{1 + n^2 x^2}$ for $n = 1, 2, 3, \ldots$ and $x \in [0,1]$. Your task is to explore whether this sequence converges to the function $f(x) = 0$ as n increases.

Hint: You might find sapply() helpful for this task. However, there are different ways to find a solution to this problem.

# a)

Write a function in R that computes $f_n(x)$ for given $n$ and $x$. Furthermore, write the target function $f(x)=0$.

```{r}

## initial function
f_n_x <- function(x, n) {
  calculation <- (n * x) / (1 + n^2 * x^2)
  return(calculation)
}


## target function f_x_0, asymp. for large n
f_x_0 <- function(x){
  return(0)
}


# Test functions
f_n_x(1, 3)
f_x_0(28)
```

Choose a few fixed points $x = 0.1, 0.3, 0.5, 0.7, 0.9, 1.0$ and values for $n = c(1:100)$ create a single plot showing the values of $f_n(x)$ for the different fixed points as $n$ increases. Add a meaningful tilte, labels for the y- and x-axis as well as a legend to the plot.

For which type of convergence are you testing with this task? Does convergence for this concept hold?

```{r}
# set x and n values
x_values <- c(0.1, 0.3, 0.5, 0.7, 0.9, 1.0)
n_sequence <- seq(from = 1, to = 100, length.out = 100)


# calculate results for all "cases" using sapply()-function, basically iterating thru values (nested loop)
results <- sapply(x_values, function(x) {
  sapply(n_sequence, function(n) {
    f_n_x(x, n)
  })
})


# plot results

max_row_index <- which.max(results[, 1])
max_row <- results[max_row_index, ]
print(max_row_index)

# Answer here
# Its called pointwise convergences.
# yes


```

```{r}

# Fixed points x
x_points <- c(0.1, 0.3, 0.5, 0.7, 0.9, 1.0)

# Sequence of n values from 1 to 100
n_values <- 1:100

# Compute the values of f_n(x) for each fixed x and n
results <- sapply(x_points, function(x) sapply(n_values, function(n) f_n(x, n)))

# Plot the results
plot(n_values, results[,1], type = 'l', col = 1, lty = 1, ylim = c(0, 1),
     xlab = "n", ylab = expression(f[n](x)),
     main = expression(paste("Values of ", f[n](x), " for different fixed x as n increases")))

# Add lines for the other x points
for (i in 2:length(x_points)) {
  lines(n_values, results[,i], col = i, lty = i)
}

# Add a legend
legend("topright", legend = paste("x =", x_points), col = 1:length(x_points), lty = 1:length(x_points))



# Answer here
# pointwise convergences.
# 

```

# b)

Now you are supposed to create a sequence of plots showing $f_n(x)$ for $n = 10, 20, 50, 100, 200$ over the interval $x \in [0, 1]$ in one single figure. Add the target function $f(x) = 0$ to each plot. Add a meaningful tilte, labels for the y- and x-axis as well as a legend to the plots.

For which type of convergence are you testing with this task? Does convergence for this concept hold?

Hint: You may use par(mfrow=c(3,2)) to create the figure.

```{r}
# Set up the multi-panel plot
par(mfrow = c(3, 2), mar = c(4, 4, 2, 1)) # 3 rows, 2 columns

# Loop through each n value to create plots
for (n in n_values) {
  # Compute f_n(x) for each x value
  fn_values <- sapply(x, function(xi) f_n_x(xi, n))
  
  # Plot f_n(x)
  plot(x, fn_values, type = 'l', col = 'blue',
       xlab = "x", ylab = expression(f[n](x)),
       main = paste("f_n(x) for n =", n),
       ylim = c(-0.5, 0.5))
  
  # Add the target function f(x) = 0
  abline(h = 0, col = 'red', lty = 2, lwd = 2)
  
  # Add a legend
  legend("topright", legend = c(expression(f[n](x)), "f(x) = 0"),
         col = c('blue', 'red'), lty = c(1, 2), lwd = 2)
}


# Answer here:
# 
```

Choose an $\epsilon > 0$, for instance, $\epsilon = 0.1$. Plot $|f_n(x) - f(x)|$ for $n = 1, 2, 5, 10, 20$ and show that there are always some values of $x$ in $[0, 1]$ such that $|f_n(x) - f(x)| > \epsilon$ to support your argument in (b).

What do the plots show and what can we conclude from this observation?

```{r}

# Set up the plotting area
par(mfrow = c(3, 2))

# Define the range of x
x <- seq(0, 1, length.out = 1000)

# Define the values of n
n_values <- c(10, 20, 50, 100, 200)

# Plot f_n(x) and f(x) for each value of n
for (n in n_values) {
  y <- f_n(n, x)
  plot(x, y, type = "l", col = "blue", ylim = c(-0.5, 0.5),
       main = paste("f_n(x) for n =", n),
       ylab = expression(f[n](x)), xlab = "x")
  abline(h = 0, col = "red", lty = 2) # Add the target function f(x) = 0
  legend("topright", legend = c(expression(f[n](x)), expression(f(x) == 0)), 
         col = c("blue", "red"), lty = c(1, 2))
}





#Answer here:
#
```

### Task 2 \### (3 Points)

Calculate Power for a permutation test

For this problem, the following simple regression model is given $y_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \varepsilon_i$, where $\quad i = 1, 2, \ldots, n$, $X_1 \sim \mathcal{U}(1,10)$, $X_2 \overset{\text{iid}}{\sim} \text{Bernoulli}(p)$, and $\varepsilon \sim \mathcal{N}(0,1)$. Further, you can assume that $y_i = 1 + 0.2 x_{1i} + 0.1 x_{2i} + \varepsilon_i$.

Whenever using random numbers please use seed = 100.

# a) Create Data

Write two functions. The first should set up the model. The second should estimates the coeffcient $\beta_1$ and its t-value. You may use the $lm()$ function for this step.

```{r}

# Set seed for reproducibility
set.seed(100)

# Function to generate data
generate_data <- function(n) {
  # Generate predictors
  X1 <- runif(n, 1, 10)  # X1 ~ U(1, 10)
  X2 <- rbinom(n, 1, 0.5)  # X2 ~ Bernoulli(0.5)
  
  # Generate error term
  epsilon <- rnorm(n, 0, 1)  # epsilon ~ N(0, 1)
  
  # Calculate response variable
  y <- 1 + 0.2 * X1 + 0.1 * X2 + epsilon
  
  # Return the data as a data frame
  return(data.frame(y = y, X1 = X1, X2 = X2))
}

# Example of generating data with n = 100
data <- generate_data(100)
head(data)

```

# b) Perform a permutation test

Perform a permutation test for the effect of $X_1$ on $Y$. First, generate data with the function you just created. Then, randomly sample $Y$ 30 times and fit the model for each sampled dataset. For each sampled dataset, calculate the coefficient for $\beta_1$ and its t-value. Calculate the p-value as the proportion of times the absolute t-values from the sampled data are greater than or equal to the absolute t-value from the original data. Repeat these steps 100 times and calculate the share of rejections at the 5% level. Perform this process for different sample sizes, starting at 50 and increasing by 50 until reaching 500.

```{r}

# Set seed for reproducibility
set.seed(100)

# Function to perform permutation test
permutation_test <- function(data, num_permutations = 100) {
  # Fit the original model
  original_model <- lm(y ~ X1 + X2, data = data)
  original_summary <- summary(original_model)
  original_t_value <- original_summary$coefficients["X1", "t value"]
  
  # Perform permutations
  perm_t_values <- numeric(num_permutations)
  for (i in 1:num_permutations) {
    # Shuffle the response variable
    permuted_y <- sample(data$y)
    
    # Fit the model to the permuted data
    permuted_model <- lm(permuted_y ~ data$X1 + data$X2)
    permuted_summary <- summary(permuted_model)
    
    # Store the t-value for X1
    perm_t_values[i] <- permuted_summary$coefficients["data$X1", "t value"]
  }
  
  # Calculate the p-value
  p_value <- mean(abs(perm_t_values) >= abs(original_t_value))
  
  return(p_value)
}

# Function to calculate power for different sample sizes
calculate_power <- function(sample_sizes, num_iterations = 100, num_permutations = 30) {
  power_results <- numeric(length(sample_sizes))
  
  for (j in seq_along(sample_sizes)) {
    n <- sample_sizes[j]
    rejections <- numeric(num_iterations)
    
    for (i in 1:num_iterations) {
      # Generate data
      data <- generate_data(n)
      
      # Perform permutation test
      p_value <- permutation_test(data, num_permutations)
      
      # Check if p-value is less than 0.05
      rejections[i] <- (p_value < 0.05)
    }
    
    # Calculate the power
    power_results[j] <- mean(rejections)
  }
  
  return(power_results)
}

# Define sample sizes
sample_sizes <- seq(50, 500, by = 50)

# Calculate power for the given sample sizes
power_results <- calculate_power(sample_sizes)

# Print power results
data.frame(SampleSize = sample_sizes, Power = power_results)



```

# c) Compare test results

Repeat the steps from task b), but set $beta_1 = 0$ for the data generation. Plot the results for the different samples sizes in the same plot as the results for taks b) and compare. Interpret the plot.

```{r}

# Function to generate data with beta1 = 0
generate_data_beta1_0 <- function(n) {
  X1 <- runif(n, 1, 10)  # X1 ~ U(1, 10)
  X2 <- rbinom(n, 1, 0.5)  # X2 ~ Bernoulli(0.5)
  epsilon <- rnorm(n, 0, 1)  # epsilon ~ N(0, 1)
  y <- 1 + 0 * X1 + 0.1 * X2 + epsilon
  return(data.frame(y = y, X1 = X1, X2 = X2))
}

# Function to calculate power for different sample sizes with beta1 = 0
calculate_power_beta1_0 <- function(sample_sizes, num_iterations = 100, num_permutations = 30) {
  power_results <- numeric(length(sample_sizes))
  
  for (j in seq_along(sample_sizes)) {
    n <- sample_sizes[j]
    rejections <- numeric(num_iterations)
    
    for (i in 1:num_iterations) {
      # Generate data with beta1 = 0
      data <- generate_data_beta1_0(n)
      
      # Perform permutation test
      p_value <- permutation_test(data, num_permutations)
      
      # Check if p-value is less than 0.05
      rejections[i] <- (p_value < 0.05)
    }
    
    # Calculate the power
    power_results[j] <- mean(rejections)
  }
  
  return(power_results)
}

# Calculate power for the given sample sizes with beta1 = 0
power_results_beta1_0 <- calculate_power_beta1_0(sample_sizes)

# Install ggplot2 package if not already installed
if (!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}

# Create data frames for plotting
power_data <- data.frame(SampleSize = rep(sample_sizes, 2),
                         Power = c(power_results, power_results_beta1_0),
                         Beta1 = factor(rep(c("0.2", "0"), each = length(sample_sizes))))

# Plot the results
ggplot(power_data, aes(x = SampleSize, y = Power, color = Beta1, group = Beta1)) +
  geom_line() +
  geom_point() +
  labs(title = "Power of Permutation Test for Different Sample Sizes",
       x = "Sample Size",
       y = "Power",
       color = "Beta1") +
  theme_minimal()

```
