---
title: "HA_Prob_Calc_Winter23_24"
author: "Rouven Lindenau"
date: "2024-01-22"
output: html_document
---
---
title: "Home Assignment, Probability Calculus 2023/2024"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Procedure
- Try to solve this home assignment and answer all questions in this file by filling the boxes below each exercise
```{r}
#Example

#Code here
```
- You are not allowed to use any additional R packages (this means packages that you have to install additionally to those already installed by default). 
- You can work in groups of up to 5 students. 
- Hand in your solution (as a Rmd file!) via [e-mail] (l.baumann@stat-econ.uni-kiel.de) and (rl@stat-econ.uni-kiel.de) until Sunday, 11.02.2024, 23:59:59. 
- Make sure to include all names and stu-numbers in your mail and in the header of this file!
- We won't evaluate your assignment if your code is not running!
- Add some comments to your code lines so people can understand your code more easily.
- We might ask you to explain your code if we suspect that you did not write it on your own.


!!! Enter your group name and stu-numbers here !!!

1. Josef Fella, stu245231
2. First name laste name, stuXXXXX 
3. First name laste name, stuXXXXX 
4. First name laste name, stuXXXXX 
5. First name laste name, stuXXXXX 

!!! Enter your group name and stu-numbers here !!!

_______________________________________________________________________________________________________________
#1. a) Exercise: Password generator (3 Points)

You are asked to write a function that creates random passwords. The function must have two arguments:
1. length - number of symbols used in the password, it should be 4 by default if no value is specified.
2. special - a TRUE/FALSE boolean that is TRUE if special characters like "$" are included, it should be TRUE by default, if no value is specified.

The password can be made of 
- lower case letters, "a", "b",...,"z" (there are 26 possibilities)
- upper case letters, "A", "B",...,"Z" (there are 26 possibilities)
- numbers, "0", "1",...,"9" (there are 10 possibilities)
- and (if allowed) special characters which are: "$", "%", "&", "/", "(", ")" (there are 6 possibilities)

Further, the password is supposed to have at least one of each: lower case letter, upper case letter, number and if allowed special character. 
Apart from that, at each position the password should have an equal probability to draw one of the 68 symbols (62 if special characters not allowed).

Additionally, 
3. If "length" in the function argument is less than 4, return: "Password must be at least of length 4!"
4. If "length" in the function argument is not an integer, return: "Number of symbols must be a positive integer!"

Last,
5. If length of the password is <6, the function should print: "Weak password!"
6. If length of the password is >6 and <10, the function should print: "Moderate password!"
7. If length of the password is >10, the function should print: "Strong password!"

8. The function should return the password as one single character, i.e. merging the drawn symbols in one character.

Make sure your function fulfills all needed conditions by testing a few variations!

```{r}

password_generator <- function(length, special) {
  # funtion to detect integers --> is.number was errours
  is.wholenumber <- function(x, tol = .Machine$double.eps^0.5) abs(x - round(x)) < tol
  
  # 1.Check: For integer and length > 4
  if (missing(length) || !is.wholenumber(length) || length < 4) {
    return("Number of symbols must be a positive integer and password must be at least of length 4!")
  }
  
  # Lists with elements for password
  letters <- strsplit("abcdefghijklmnopqrstuvwxyz", NULL)[[1]]
  letters_upper <- strsplit("ABCDEFGHIJKLMNOPQRSTUVWXYZ", NULL)[[1]]
  numbers <- strsplit("0123456789", NULL)[[1]]
  specials <- strsplit("$%&/()", NULL)[[1]]
  
  # 2. Check: If specials included or not
  lst <- c(letters, letters_upper, numbers)
  if (special) {
    lst <- c(lst, specials)
  }
  
  # Creates an output vector and adds a random element from each list
  output <- c(
    sample(letters, 1),
    sample(letters_upper, 1),
    sample(numbers, 1)
  )

  if (special) {
    output <- c(output, sample(specials, 1))
  }
  
  # Loop thu list until desired length is fulfilled
  while (length(output) < length) {
    random_element_selector <- sample(lst, 1)
    output <- c(output, sample(random_element_selector, 1))
  }
  
  # Setting up final password
  final_password <- paste(sample(output), collapse = "")
  
  # Output of the password and message for user
  if (nchar(final_password) < 6) {
    return(paste("Weak password!", final_password))
  } else if (nchar(final_password) >= 6 && nchar(final_password) < 10) {
    return(paste("Moderate password!", final_password))
  } else {
    return(paste("Strong password!", final_password))
  }
}

```

#1. b) Test your function here for

```{r}
# Test of the function

length = 7
special = T
password_generator(length, special)


length = 3
special = T
password_generator(length, special)


length = 4.5
special = T
password_generator(length, special)


length = 5
special = F
password_generator(length, special)


length = 50
special = T
password_generator(length, special)

```


# 2. a) Excercise: Gamma vs. Normal distribution (3 points):

In this exercise you need to demonstrate that if a random variable $X$ is ${\Gamma}$-distributed with shape parameter $a$ and scale parameter $b$, then for $a \to \infty$  $X \overset{d}{\rightarrow} \mathcal{N} (ab,ab^2)$ if $a > b$.

To this end, write a function that samples and returns $N$ observations from the ${\Gamma}$ distribution with with shape parameter $a$ and scale parameter $b$. The function's input shall be $a$, $b$ and N.

```{r}
# Setting fixed random seed
set.seed(28)

# Function to get gamma sample (Inverse Transformation method)
get_gamma_sample <- function(a, b, N) {
  # Generate uniform random variables
  u <- runif(N)
  
  # Calculate the scale parameter for qgamma
  scale_param <- 1 / b
  
  # Use qgamma to get Gamma-distributed sample
  gamma_sample <- qgamma(u, shape = a, scale = scale_param)
  return(gamma_sample)
}

# Generate gamma-distributed samples for large alpha
alpha_large <- 5
beta <- 1
N <- 1000

gamma_samples_large_alpha <- get_gamma_sample(alpha_large, beta, N)

# Plot histogram
hist(gamma_samples_large_alpha, freq = FALSE, main = "Gamma vs Normal Distribution",
     xlab = "Value", col = "lightblue", border = "black")

# Add the theoretical normal distribution curve
curve(dnorm(x, mean = alpha_large * beta, sd = sqrt(alpha_large * beta^2)), 
      col = "red", lwd = 2, add = TRUE)

# Add vertical lines for means
abline(v = mean(gamma_samples_large_alpha), col = "blue", lty = 2, lwd = 2)
abline(v = alpha_large * beta, col = "green", lty = 2, lwd = 2)

# Add legend with smaller size
legend("topright", legend = c("Gamma Dist", "Normal Dist", "Gamma Mean", "Normal Mean"),
       col = c("lightblue", "red", "blue", "green"), lty = c(1, 1, 2, 2), lwd = c(1, 2, 2, 2), cex = 0.8)


```

Plot the resulting observations from the ${\Gamma}$ distribution using histograms with Scott breaks for $a$ = c(1, 5, 10, 50, 100) and $b$ = 1, N = 5000 in a 3x2 plot window and add the respective limiting Normal density for comparison. Choose a meaningful title, label the axes and add a legend to your plot.

```{r}
# Setting parameters
alpha_values <- c(1, 5, 10, 50, 100)
beta <- 1
N <- 5000
set.seed(28)

# Setting up subplots with larger size and more elongated aspect ratio
par(mfrow = c(3, 2), figsize = c(12, 30), mar = c(4, 4, 2, 2))

# Creating plots
for (i in 1:length(alpha_values)) {
  
  # Generate Gamma samples
  gamma_samples <- get_gamma_sample(alpha_values[i], beta, N)
  
  # Plot Gamma distribution histogram
  hist(gamma_samples, breaks = "scott", col = 'blue', main = paste("Gamma Distribution, Alpha =", alpha_values[i], ", Beta =", beta),
       xlab = "Value", ylab = "Density", xlim = c(min(gamma_samples), max(gamma_samples)))
  
  # Plot the limiting Normal distribution
  normal_mean <- alpha_values[i] * beta
  normal_var <- alpha_values[i] * (beta^2)
  normal_samples <- rnorm(N, mean = normal_mean, sd = sqrt(normal_var))
  hist(normal_samples, breaks = "scott", col = 'red', add = TRUE, density = 10)
  
  # Add legend with adjusted cex value for better readability
  legend("topright", legend = c('Gamma Distribution', 'Normal Distribution (Limit)'), fill = c('blue', 'red'), cex = 0.8)
}

```

# 2. b) Slutsky's theorems

Now assume that you observe $X_n \sim \Gamma(a,b)$ for $a = c(1, 5, 10, 50, 100)$ and $b = 1$ as well as $Y_n \sim \mathcal{N}\left(\mu, 4\sigma^2\sqrt{n^{-5}}\right)$ where $\mu = 3.5$ and $\sigma = 7$. Verify that following part of Slutsky's theorems holds with increasing values of $a$ : 2. $X_n \cdot Y_n \overset{d}{\rightarrow} X\cdot c$. Do this with the help of Monte Carlo simulations and a corresponding 3x2 plot window. Choose $N_{MC} = 1000$, $n = 1000$.
Hint: Remember that $X \overset{d}{\rightarrow} \mathcal{N} (ab,ab^2)$ for increasing values of $a$.

```{r}
# Empirical work on Slutsky's theorem

# Set parameters
N_MCarlo <- 1000
n <- 1000
alpha_values <- c(1, 5, 10, 50, 100)
beta <- 1
mu <- 3.5
sigma <- 7

set.seed(28)

# Set up subplots with larger size and elongated aspect ratio
par(mfrow = c(3, 2), mar = c(4, 4, 2, 2), fig.width = 10, fig.height = 12)

# Create plot
for (i in 1:length(alpha_values)) {
  # Generate sample
  X_n <- get_gamma_sample(alpha_values[i], beta, N_MCarlo)
  Y_n <- rnorm(N_MCarlo, mean = mu, sd = 4 * (sigma^2) * sqrt(n^-5))
  sample <- X_n * Y_n

  # Plot results
  hist(sample, breaks = "scott", prob = TRUE, col = 'blue',
       main = paste("Gamma * Normal, Alpha =", alpha_values[i], ", Beta =", beta),
       xlab = "Value", ylab = "Density", xlim = c(min(sample), max(sample)))
  
  # Compute expected limiting distribution parameters
  expected_mean <- alpha_values[i] * mu
  expected_variance <- alpha_values[i] * (mu^2 + 16 * (sigma^4) * (n^-5))

  # Overlay expected limiting distribution without additional libraries
  x <- seq(expected_mean - 3 * sqrt(expected_variance), expected_mean + 3 * sqrt(expected_variance), length.out = 100)
  pdf_values <- (1 / sqrt(2 * pi * expected_variance)) * exp(-0.5 * ((x - expected_mean) / sqrt(expected_variance))^2)
  lines(x, pdf_values, col = 'red', lwd = 2)
  
  legend("topright", legend = c('Empirical', 'Limiting Distribution'), fill = c('blue', 'red'))
}

```